---
title: "PsyR"
output: html_document
date: "2024-11-06"
---

## This R package is designed to replicate the result from Psy, which is a software that constructs confidence intervals on contrasts in ANOVA designs. 

Load the necessary packages
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(emmeans)
library(afex)
library(readxl)
library(PsyR)
```

### This example data is mean metabolic content for 7 brain regions in 74 healthy participants. We want to to compare each region to each other region, so this is a repeated measures design with one within-subject factor

Read the data
```{r}
data <- read_excel("../Desktop/mrs/new ratios new masks/74_PME_to_pooled.xlsx")
```




# Psy output, which we aim to replicate

For this list of contrasts

              1 -1 0 0 0 0 0
              1 0 -1 0 0 0 0
              1 0 0 -1 0 0 0

Psy outputs these confidence intervals, when using the post hoc method 

                            Raw CIs (scaled in Dependent Variable units)
              -------------------------------------------------------
              Contrast      Value        SE           ..CI limits..  
                                                    Lower       Upper
              -------------------------------------------------------
              W1            0.041       0.002       0.037       0.046
              W2           -0.005       0.001      -0.007      -0.002
              W3            0.008       0.001       0.004       0.012
              -------------------------------------------------------



# The traditional route - ANOVA, emmeans, Bonferroi
### Specify a multivariate model for emmeans since repeated measures are present.
```{r}
afex_options(emmeans_model = "multivariate")
```

### Fit the ANOVA model with a repeated measures design with one within-subject factor (region)
```{r}
mod <- aov_ez("subject","content", data, within = "region")
mod
```

### Compute estimated marginal means (EMMs)
```{r}
em_results <- emmeans(mod, ~ region)
em_results
```

### Get a contrast table for pairwise comparisons, using Bonferroni correction
```{r}
pairwise_comparisons <- contrast(em_results, method = "pairwise", adjust = "bonferroni")
pairwise_comparisons
```

### Compute confidence intervals for these contrasts
```{r}
confint(pairwise_comparisons)
```

##The confidence intervals don't match the Psy output.
For example, the confidence interval for basal_ganglia - cerebellum is 0.048 in Psy, whereas emmeans outputs 0.047.



# The implementation of the PsyR package on the same data
### Specify a multivariate model for emmeans since repeated measures are present.
```{r}
afex_options(emmeans_model = "multivariate")
```

### Fit the ANOVA model with a repeated measures design with one within-subject factor (region)
```{r}
mod <- aov_ez("subject","content", data, within = "region")
mod
```

### Compute estimated marginal means (EMMs)
```{r}
em_means <- emmeans(mod, "region")
em_means
```

### Get the contrast table for all regional pairwise comparisons 
```{r}
pairwise_contrasts <- contrast(em_means, method = "pairwise")
pairwise_contrasts
```

# Now we implement the PsyR package to get confidence intervals of the contrasts, using the post hoc method
```{r message=FALSE, warning=FALSE, echo=FALSE}

psyci (
  mod,
  pairwise_contrasts,
  "bf",
  family = "w",
  within_factors = list("region"),
  alpha = 0.05
)

```
